---
title: "Data Analysis Project"
author: "Ayesha Abdulllahi"
date: ""
output: html_document
---

</br/>

## ___GENERAL INSTRUCTIONS___

1. Fill in your name in the 'author' section in line 3.
2. In each code chuck, write the required code to answer the question.
3. In each answer section, write the answer to the question based on the output from the code you have developed.
4. Save the document as you proceed through the questions.
5. Email the completed assignment (Markdown File) to me by 7:20pm on 12/01 at 7:20pm.

Good luck!


</br/>

```{r echo = FALSE, warning=FALSE, message = FALSE}
# LOAD LIBRARIES
library(mlbench)
library(ggplot2)
library(GGally)
library(corrplot)
library(expss)
library(tidyverse)
library(stringr)
library(leaflet)
library(leaflet.extras)
library(dplyr)
library(scales)

```


</br/>


## ___US ARRESTS DATA___

In this section use the "USArrests" data.  This dataset provides state level information on the murder rate, assault rate, rape rate, and urban population.

```{r echo = FALSE}
# LOAD DATA

load("/cloud/project/USArrests.RDat")


```


</br/>


___1. What is the mean and median murder rate? (1 point)___

```{r}

mean(USArrests$Murder)

median(USArrests$Murder)


```

Answer: 
The mean murder rate is 7.788 and the median murder rate is 7.25.


</br/>


___2.  Create a dataset that only includes Virginia and surrounding states (West Virginia, Maryland, North Carolina, Tennessee, Kentucky) in descending order by Murder rate. (1 point)___
```{r}
VA_others <- USArrests  %>%  filter(State == "Virginia" | State == "West Virginia" | State == "Maryland" | State == "North Carolina" | State == "Tennessee" | State == "Kentucky") %>%
  
  select(Murder, State) %>%
   count(Murder, State) %>%     
  arrange( desc(Murder) )              
  
```


</br/>


___3. Create a table with the top five states that have the highest reported assault rate.  (1 point)___

```{r}
five_statesA <- USArrests  %>%  filter(State == "Maryland" | State == "North Carolina" | State == "Florida" | State == "Arizona" | State == "New Mexico") %>%
  
  select(State, Assault) %>%
   count(State, Assault) %>%     
    arrange( desc(Assault) )      

cro( five_statesA$Assault, five_statesA$State ) %>% tab_sort_desc()

```


</br/>


___4. Create a bar graph with: (1) state on the X axis, and (2) murder on the Y axis.  Based on the bar graph, which state had the highest rate of murder?  (1 point)___

```{r}
SYbar <- USArrests %>%           
  count(State, Murder) %>%     
  arrange( desc(Murder) )              
  

ggplot(data = SYbar, aes(x = reorder(State, Murder), y =  Murder)) +  
  geom_bar(stat="identity") +                  
  theme(axis.text.x=element_text(angle=75,      
                                 size = 8,       
                                 hjust=0.95,     
                                 vjust=0.95))     
```

Answer: 
The state with the highest rate of murder is Georgia.

</br/>
</br/>
</br/>
</br/>
</br/>


## ___CITY CRIME DATA___

In this section use the "crime_data" data.  This dataset provides basic information for crime events that occurred across several cities in 2017-2018.

```{r echo = FALSE}
# LOAD DATA

load("/cloud/project/crime_data.RDat")


```


</br/>


___5. What city has the highest number of incidents? (1 point)___
```{r}
Q5 <- crime_data  %>% 
  
  select(city_name, offense_against) %>%
   count(city_name, offense_against) %>%     
    arrange( desc(offense_against) )     
cro(crime_data$city_name, crime_data$offense_against) %>% tab_sort_desc()

```

Answer: 
New York City has the highest number of incidents!

</br/>


___6. Which city has the highest number of 'personal robbery' offenses based on the offense type variable. (1 point)___
```{r}
PRhigh <- crime_data %>% filter(offense_type == "personal robbery") %>%    
  
  select(city_name, offense_type) %>%  
  
  count(city_name, offense_type) %>%   
  
  arrange( desc(n)) %>%
  drop_na()


```

Answer: 
Chicago has the highest number of "personal robbery" offenses.

</br/>


___7. What is the most common offense type? (1 point)___

```{r}
COT <- crime_data %>% 
  
  select(offense_type) %>%  
  count(offense_type) %>%  
  arrange( desc(n)) %>%
  drop_na()


```

Answer:
The most common offense type is simple assault.

</br/>


___8. What city has the highest rate of the most common offense type identified in question 7? (1 point)___

```{r}
Q8 <- crime_data %>% filter(offense_type == "simple assault") %>% 
  
  select(city_name, offense_type) %>%  
  count(city_name, offense_type) %>%   
  arrange( desc(n)) %>%
  drop_na()


```

Answer: 
The city with the most simple assaults is Chicago.

</br/>


___9. What is the most common offense type that is located (location_type) in restaurants? (1 point)___

```{r}

locrest <- crime_data %>% filter (location_type == "restaurant") %>% 
  
  
  select(offense_type, location_type) %>%  
  count(offense_type, location_type) %>%   
  arrange( desc(n)) %>%
  drop_na()
   
```

Answer: 
the most common offense type that is located in restaurants is "theft from building".

</br/>


___10. In what location type do 'drug/narcotic violations' most frequently occur? (1 point)___

```{r}

Q10 <- crime_data %>% filter (offense_type == "drug/narcotic violations") %>% 
  
  
  select(offense_type, location_type) %>%  
  
  count(offense_type, location_type) %>% 
  arrange( desc(n)) %>%
  drop_na()

```

Answer: 
The location type where 'drug/narcotic violations' most frequently occur is the street.

</br/>


___11. What is the most common offense type in Chicago? (1 point)___

```{r}
coofchi <- crime_data %>% filter (city_name == "Chicago") %>% 
  
  
  select(offense_type, city_name) %>%  
  count(offense_type, city_name) %>%   
  arrange( desc(n)) %>%
  drop_na()

```

Answer: 
The most common offense type in Chicago is simple assault.

</br/>

___12. Create a crime map of all offenses in the 'crime_data' dataset.  What does that map show you?___

```{r}
leaflet(crime_data) %>%                               
  addTiles() %>%                                  
  addProviderTiles("CartoDB.DarkMatter") %>%      
  addCircleMarkers(
                   color = "red",
                   radius = 1,
                   stroke = FALSE, 
                   fillOpacity = 0.5)

```

Answer: 
The map shows me that crime can be high in two different cities that happen to be in the same state (i.e. California, Texas, and Arizona).

</br/>


___13. Create a crime map of personal robbery offenses in Chicago___

```{r}
MAP_prchi <- crime_data %>% filter(offense_type == "personal robbery" & city_name == "Chicago") 



leaflet(MAP_prchi) %>%                               
  addTiles() %>%                                  
  addProviderTiles("CartoDB.DarkMatter") %>%      
  addCircleMarkers(
                   color = "red",
                   radius = 1,
                   stroke = FALSE, 
                   fillOpacity = 0.5)

```


</br/>


___14. Create a heatmap of drug/narcotic violations in New York___

```{r}
heatmap <- crime_data %>% 
  filter(offense_type == "drug/narcotic violations" & city_name == "New York")  


leaflet(heatmap) %>%                    
  addProviderTiles("CartoDB.DarkMatter") %>%  
  addHeatmap(      
    blur = 20, 
    max = 2.0, 
    radius = 10)

```


</br/>
</br/>
</br/>
</br/>
</br/>


## ___FARGO___

In this section use the "FARGO" dataset.  This dataset provides information on crime incidents that occurred in Fargo in 2017.

```{r echo = FALSE}
# LOAD DATA
 load("/cloud/project/FARGO.RDat")


```


</br/>


___15. Which month had the most crime? (1 point)___

```{r}

wm <- FARGO %>% 
  
  
  select(Month, Call.Type) %>%  
  count(Month) %>%   
  arrange( desc(n)) %>%
  drop_na()

```

Answer: 
The month with the most crime is June.

</br/>


___16. At what address are the most crimes committed? (1 point)___


```{r}
WA <- FARGO %>% 
  
  
  select(Address, Call.Type) %>%  
  count(Address) %>%  
  arrange( desc(n)) %>%
  drop_na()


```

Answer: 
Most crimes are committed at 200 BLK 4 ST N Fargo, ND 58102.

</br/>


___17. What is the most frequent type of crime (based on call type) that happens at the address identified in question 16? (1 point)___

```{r}

Q17 <- FARGO %>% filter(Address == "200 BLK 4 ST N Fargo, ND 58102") %>%
  
  
  select(Address, Call.Type) %>% 
  count(Address, Call.Type) %>%  
  arrange( desc(n)) %>%
  drop_na()

```

Answer: 
The most frequent type of crime that happens at 200 BLK 4 ST N Fargo, ND 58102 is harassment.

</br/>


___18. In March, on what day of the week do most crimes occur? (1 point)___

```{r}

March <- FARGO  %>% filter(Month == "Mar" ) %>%

 group_by(Weekday) %>%   

 select(Month, Weekday, Call.Type) %>%  

 count(Weekday) %>%  

 arrange(desc(n)) %>%     

 drop_na() 

```

Answer: 
In March, crime is mostly occurring on Fridays. 

</br/>


___19. On what day of the week are burglaries most likely to occur? (1 point)___

```{r}
Burglaries <- FARGO  %>% filter(Call.Type == "Burglary" ) %>%

 group_by(Weekday) %>%   

 select(Weekday, Call.Type) %>%  

 count(Weekday) %>%

 arrange(desc(n)) %>%     

 drop_na() 


```

Answer: 
Burglaries are most likely to occur on Saturdays.

</br/>


___20. Between June-August, what day of the week are assaults most likely to occur? (1 point)___

```{r}


June_Aug <- FARGO  %>% filter(Month == "June"| Month == "July"| Month ==  "August"| Call.Type == "Assault" ) %>%

 group_by(Weekday) %>%   

 select(Month, Weekday, Call.Type) %>%  

 count(Weekday) %>%  

 arrange(desc(n)) %>%     

 drop_na() 
```

Answer: 
Between June-August, assaults are most likely to occur on Saturdays.







